Music Versus No Music Effectiveness On Cognitive Response Time and Typing Efficiency

### 2024-04-30 paper/paper.tex

\section{Related Works}
While there is not necessarily a concrete answer for whether music or background noise is effective on cognitive function, there have been many
studies that attempt to determine the answer. The other studies attempted to search for this answer by
focusing on word processing on a computer with the background music and the user's ability to formulate sentences~\cite{ransdell2001141}, 
or determining if background music helps users to keep their attention on their task at hand~\cite{kiss2021effect}. Similar to~\cite{kiss2021effect},~\cite{mathew2022inherently} also examined how motivation and music were used hand-in-hand on ``a time-consuming task and task performance''. A different study done in 2010 again found similar results of music both being a hindrance and benefit~\cite{AudioDistractionsAshley}. According to~\cite{Gonzalez_Aiello_2019}, the degree of cognitive function may be based on an individual's personality.

Although~\cite{ransdell2001141} suggests that the music is a detriment, those with ``musical training and high working memory span wrote better essays with longer sentences'. This suggests that while the music slowed down the participants, it did not necessarily lower their cognitive ability in typing on their computer while focusing on forming text for the written essay for this particular study. For~\cite{AudioDistractionsAshley}, the music caused enough of a difference in the time that was taken for writing a paragraph but ``there were no significant differences in typing speed between the audio distractions''. This again may suggest that the participant's cognitive ability was not hindered when typing and may transfer directly after should the participants have taken a test measuring their cognitive function. The slow timing was again seen in~\cite{Gonzalez_Aiello_2019} in which ``music generally impaired performance on a complex task but improved performance on a simple task''.

The study done through~\cite{kiss2021effect} saw opposite results as to~\cite{Gonzalez_Aiello_2019}. The music or background noise ``enhanced'' the reaction time and performance of the tasks that are given to be completed. The results from~\cite{mathew2022inherently} were split. The music was a detriment for logic puzzles in the user's accuracy but aided in the efficiency of a writing task.

>>>>

\section{Related Works}
While there is not necessarily a concrete answer for whether music or background noise is effective on cognitive function, there have been many
studies that attempt to determine the answer. 
The other studies attempted to search for this answer by focusing on word processing on a computer with the background music and the user's ability to formulate sentences~\cite{ransdell2001141}, or determining if background music helps users to keep their attention on their task at hand~\cite{kiss2021effect}. 
Similar to Kiss and Linnell~\cite{kiss2021effect}, Mathew ~\cite{mathew2022inherently} also examined how motivation and music were used hand-in-hand on ``a time-consuming task and task performance''. 
A different study done in 2010 again found similar results of music both being a hindrance and benefit~\cite{AudioDistractionsAshley}. 
According to Gonzalez and Aiello~\cite{Gonzalez_Aiello_2019}, the degree of cognitive function may be based on an individual's personality. 
Furnham et al., also determined this was based on personality in a study done between extraverts and introverts although ``no significant interactions were found''~\cite{distractingEffectsFurnham}.

The research done by Komlao found that music or background music was again based on an individual's personality but that music, even music that the participants were familiar with, were a distraction and ``was found to negatively impact task performace'' overall~\cite{komlao2018variations}.
This was contrary to the previous studies found by Komlao where the ``music could help increase productivity for simple, monotonouse tasks''~\cite{komlao2018variations}.

Although Ransdell and Gilroy~\cite{ransdell2001141} suggests that the music is a detriment, those with ``musical training and high working memory span wrote better essays with longer sentences''. 
This suggests that while the music slowed down the participants, it did not necessarily lower their cognitive ability in typing on their computer while focusing on forming text for the written essay for this particular study. 
For Silverberg~\cite{AudioDistractionsAshley}, the music caused enough of a difference in the time that was taken for writing a paragraph but ``there were no significant differences in typing speed between the audio distractions''. 
This again may suggest that the participant's cognitive ability was not hindered when typing and may transfer directly after should the participants have taken a test measuring their cognitive function. 
The slow timing was again seen by Gonzalez and Aiello~\cite{Gonzalez_Aiello_2019} in which ``music generally impaired performance on a complex task but improved performance on a simple task''.

The study done through Kiss and Linnell~\cite{kiss2021effect} saw opposite results as to Gonzalez and Aiello~\cite{Gonzalez_Aiello_2019}. The music or background noise ``enhanced'' the reaction time and performance of the tasks that are given to be completed. 
The results from Mathew~\cite{mathew2022inherently} were split. 
The music was a detriment for logic puzzles in the user's accuracy but aided in the efficiency of a writing task. 
According to Bramwell-Dicks et al., the instrumental music used in their study had better results on typists in speed and accuracy whereas music with vocals, the participants fared worse~\cite{bramwell2016can}.
While only a quasi-experiment, Oldham et al., found that personal headphones and Walkmans for employees had a postive correlation if the job complexity was not complex but also that employee satisfaction at this company was higher if they were allowed to use their walkman while working~\cite{oldhamListen}. Dobbs et al., also found the correlation that while music produced lower scores than silence it was better than just ``noise''~\cite{effectsOfBackgroundDobbs}.
